Comparison of Spatial Partitioning Strategies

### 2025-01-23 paper/paper.tex
Paper: Material and Method and start writing Result.

>>>>

\section{Material and Method}

\subsection{Linux}
Linux is the operating system used to develop the application and write this
paper. The program uses a specific Linux API to acquire the current time as a
UNIX time stamp. The way to do this would differ on other operating systems.

\subsection{Neovim}
Neovim was the editor used to edit the code. It is a modal text editor and not a
full integrated development environment, or IDE as they are called. A plugin
which utilizes the language server protocol was used to provide me with
intellisense and to aid in development by warning me about syntax errors.

\subsection{Git}
Git is a version control system which helps track changes to the project. In
this project it is utilized to keep a time line of the development process and
make sure changes could be synced between devices. The local git repository is
mirrored on GitHub making it accessible from other devices, and the source code
can be read from the browser.

\subsection{CMake}
CMake is a buildsystem, used to generate generate the final executable from the
source code. It aids in compiling and linking the program making cross
compilation and dependency management trivial.

\subsection{C}
C was the program of choice for this project due to its low level control.
Since memory is managed manually in C I could be sure the results would not be
skewed by a garbage collector running in the middle of a benchmark.

\subsection{SDL2}
SDL2 - Simple DirectMedia Layer - is a cross-platform library supplying an easy
API to create a window and render to it. This was crucial when developing since
visualizing the data structures made it easier to spot errors than simply
looking at the code. It also gives a nice visual representation of the program.

\subsection{Python}
Python is an interpreted programming language which was used for making a script
to generate graphs. All the graphs used in this paper was generated from data
collected from a C program, which dumped said data into a JSON file, which was
then read by a Python script using the matplotlib library to generate nice
looking graphs.

\section{Results}
This section presents the results of the benchmarking done on the three
strategies - quadtrees, spatial hashing and static grids. The metrics measured
was average runtime: the sum of all operations needed to run the program.
Insertions: how long it took to insert all the objects into the data structure.
Collision testing: total time spent checking objects against each other, which
includes querying. Querying: average time spent retrieving objects from the data
structure. Clearing: how long it takes to reset the data structure making it
ready for rebuilding. Each metric was measured with a range of object count,
labeled as 'box count' in the graphs, from ten to 2500 with an increase of ten
objects between each measurement. At each stage was run 32 times and the results
were averaged to give smoother graphs. Two benchmarks was run, one with an even
distribution of objects and one with an uneven distribution.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{even-naive.png}
    \caption{Runtime of naive solution and partitioning strategies.}
    \label{fig:naive}
\end{figure}

Figure \ref{fig:naive} shows a drastic difference in runtime between the naive
solution and the optimization strategies. It is hard to evaluate which strategy
performs best due to the stark difference between the naive and the optimized
versions. Due to this discrepancy the naive solution will not be included in the
other graphs presented. This comparison was only run with 1000 objects due to
the time taken per execution.

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=0.75\textwidth]{even-average.png}
%     \caption{Average runtime of spatial partitioning strategies.}
%     \label{fig:even-average}
% \end{figure}
